### 2024-10-24 Article/Appendix/Omitted_Proofs.tex
Reordered files + new version EqGrph

\chapter{Omitted Proofs}

\paragraph{\Cref{lemma:pullback_lemma}.}
    Suppose that the following diagram is given and its right half is a pullback. Then the whole rectangle is a pullback if and only if its left half is a pullback.
    \[
        \begin{tikzcd}
            X \arrow[r, "f"] \arrow[d, "t"swap] & Y \arrow[r, "g"] \arrow[d, "k"] & Z \arrow[d, "h"] \\
            A \arrow[r, "a"swap] & B \arrow[r, "b"swap] & C
        \end{tikzcd}
    \]
\begin{proof}
    For the ``only if'' part, suppose the left square to be a pullback. To verify the outer rectangle is a pullback, consider the following situation:
    \[
        \begin{tikzcd}
            P \arrow[bend left=30]{rrrd}{p_1} \arrow[bend right=30, swap]{ddr}{p_2} & & \\
            & & & Z \arrow[d, "h"] \\
            & A \arrow[r, "a"swap] & B \arrow[r, "b"swap] & C
        \end{tikzcd}
    \]
    But the right square is a pullback implies that there exists a unique $u$ such that
    \[
        \begin{tikzcd}
            P \arrow[bend left=30]{rrrd}{p_1} \arrow[bend right=30, swap]{ddr}{p_2} \arrow[bend left=15, dashed]{drr}{u}& & \\
            & &  Y \arrow[r, "g"] \arrow[d, "k"]& Z \arrow[d, "h"] \\
            & A \arrow[r, "a"swap] & B \arrow[r, "b"swap] & C
        \end{tikzcd}
    \]
    And, since the left square is a pullback, there exists a unique $v$ such that
    \[
        \begin{tikzcd}
            P \arrow[bend left=30]{rrrd}{p_1} \arrow[bend right=30, swap]{ddr}{p_2} \arrow[bend left=15]{drr}{u} \arrow[dr, "v", dashed]& & \\
            & X \arrow[r, "f"] \arrow[d, "t"swap] &  Y \arrow[r, "g"] \arrow[d, "k"]& Z \arrow[d, "h"] \\
            & A \arrow[r, "a"swap] & B \arrow[r, "b"swap] & C
        \end{tikzcd}
    \]
    Hence, the whole rectangle is a pullback.

    For the ``if'' part, consider the following situation.
    \[
        \begin{tikzcd}
            P \arrow[bend left=30]{rrd}{p_1} \arrow[bend right=30, swap]{ddr}{p_2} & \\
            & & Y \arrow[d, "k"] \\
            & B \arrow[r, "b"swap] & C
        \end{tikzcd}
    \]
    We have now to show that the unique arrow $v: P \to X$ (of the outer rectangle) is such that $f \circ v = p_1$, but this follows from the fact that the right square is a pullback, having the following situation.
    \[
        \begin{tikzcd}
            P \arrow[bend left=30]{rrrd}{g \circ p_1} \arrow[bend right=30, swap]{ddr}{p_2} \arrow[bend left=15, dashed]{drr}{u}& & \\
            & &  Y \arrow[r, "g"] \arrow[d, "k"]& Z \arrow[d, "h"] \\
            & A \arrow[r, "a"swap] & B \arrow[r, "b"swap] & C
        \end{tikzcd}
    \]
    \todo{Rivedere questa dimostrazione}
\end{proof}

\paragraph{\Cref{th:limit}.}
 Let $\cat C$ be a category. Then $\cat C$ has all finite limits if and only if $\cat C$ has all finite products and all equalizers.
\begin{proof}
    Let $D: \cat{I \rightarrow C}$ a diagram, with $\cat I$ finite.

    The \emph{if} statement follows from definitions of products and equalizers (\Cref{def:pullback_pushout}, \Cref{def:equalizer_coequalizer})

    To satisfy the \emph{only if} statement, we want an object $L$ together with morphisms $p_i : L \rightarrow D(j)$ such that:
    \begin{enumerate}
        \item\label{item:cone} $\{p_i: L \rightarrow D(i)\}$ is a cone -- i.e., for each morphism of $\cat I$ $\alpha : i \rightarrow j$, $D(\alpha) \circ p_i = p_j$; and
        \item\label{item:univ_prop} for each $E$ and $q_i : E \rightarrow D(j)$ in $\cat C$, with $D(\alpha) \circ q_i = q_j$ for each $\alpha : i \rightarrow j$ of $\cat I$, there exists a unique $f: E \rightarrow L$ such that $q_i = p_i \circ f$ for each $i \in \Ob(\cat I)$.
    \end{enumerate}

    Consider the two products (which exist by hypothesis) $\prod_{j \in \Ob(\cat I)} D(j)$, the product of the objects of the diagram, and $\prod_{\alpha \in \Hom(\cat I)}D(cod \ \alpha)$, the product of the codomains of the morphisms in $D$, where $\pi_x$ is the $x$-th projection of the product.
    Let now:
    \[
        \gamma, \varepsilon : \prod_{j \in \Ob(\cat I)}D(j) \ \longrightarrow \prod_{\alpha \in \Hom(\cat I)} D(cod \ \alpha)
    \]
    be defined by $\gamma_\alpha = \pi_{D(cod\ \alpha)}$ (the projection on the codomain of $\alpha$) and $\varepsilon_\alpha = D(\alpha) \circ \pi_{D(dom \ \alpha)}$.
    Let now $e: L \rightarrow \prod_{j \in \Ob(\cat I)}D(j)$ the equalizer of $\gamma$ and $\varepsilon$ (which exists by hypothesis), and, for each $j \in \Ob(\cat I)$, $p_j: L \rightarrow D(j)$, defined by $p_j = \pi_{D(j)} \circ e$.

    What we want now is to show that $(L, (p_i))_{i \in \cat I}$ is the limit of $D$, namely, to prove that the conditions given at the beginning are valid.

    For condition~\ref{item:cone}, we have to show that, for each $\alpha : i \rightarrow j$ of $\cat I$, we have $D(\alpha) \circ p_i = p_j$:
    \begin{align*}
        D(\alpha) \circ p_i 
            &= D(\alpha) \circ \pi_{D(i)} \circ e   && \textit{Definition of $p_j$} \\
            &= \varepsilon_{\alpha} \circ e         && \textit{Definition of $\varepsilon$}\\
            &= \gamma_\alpha \circ e                && \textit{$e$ is an equalizer of $\pi, \varepsilon$}\\
            &= \pi_{D(j)} \circ e                   && \textit{Definition of $\pi$} \\
            &= p_j                                  && \textit{Definition of $p_j$} \\
    \end{align*}

    For condition~\ref{item:univ_prop}, suppose that $(E, (q_i)_{i \in \Ob(\cat I)})$ has the properties stated. By definition of product, there exists a (unique) arrow $q: E \rightarrow \prod_{j \in \Ob(\cat I)}D(j)$. For each arrow $\alpha: i \rightarrow j$, we have:
    \begin{align*}
            \gamma_\alpha \circ q 
                &= \pi_{D(j)} \circ q           && \textit{Definition of $\pi$} \\
                &= q_j                          && \textit{Definition of $q_j$} \\
                &= D(\alpha) \circ q_i          && \textit{Assumption on $q_j$} \\
                &= D(\alpha) \circ \pi_{D(j)} \circ q   &&\textit{Definition of $q_i$} \\
                &= \varepsilon_{\alpha} \circ q && \textit{Definition of $\varepsilon$} \\
    \end{align*}
    Since $e$ equalizes $\pi$ and $\varepsilon$, there exists a unique $f: E \rightarrow L$ in $\cat C$ such that $q = e \circ f$. Then, for each $j \in \Ob(\cat I)$, we have $\pi_{D(j)} \circ q = \pi_{D(j)} \circ e \circ f$, hence, $q_i = p_i \circ f$.
\end{proof}