### 2024-11-21 Article/Conclusions.tex
Intro, Conc

\chapter*{Conclusions}
\addcontentsline{toc}{chapter}{Conclusions}

The result we ended up with is then $\Reg(\EGG))$-adhesivity, where $\EGG$ is a category of functors which have $\Set$ as codomain category.
The main question is now, is it strictly necessary to have $\Set$ as codomain of this functors? The entire work is developed in a more general perspective, using the least possible concept directly connected with set theory, using instead categorial constructions. The property of $\Set$ we have made large use of is the equivalence between quotient and product subobjects, which holds in exact categories.
Hence, this work can be viewed as a starting point to a more general study on e-graphs, in which edges, nodes and equivalence classes can be not only sets but objects of any exact category.