### 2025-01-29 icgt/fact_sys.tex
Fact_Sys

\subsection{Factorization systems}

\begin{definition}[Factorization System]\label{def:fact_sys}
	Let $\X$ be a category, and let $\E$, $\M$ be two classes of arrows.
	Then, $(\E, \M)$ is a \emph{factorization system} if:
	\begin{enumerate}
		\item $\E$ and $\M$ are closed under composition with isomorphisms: if $f: X \to Y$ belongs to $\E$ (to $\M$), and $h: Y\to Z$ is an isomorphism, then $h \circ f$ belongs to $\E$ (to $\M$);
		\item every arrow $f: X \to Y$ of $\X$ admits an \emph{$(\E, \M)$-factorization}, that is, there are arrows $e \in \E$ and $m \in \M$ such that $f = m\circ e$;
		\item every $e \in \E$ has the \emph{left lifting property} with respect to every $m \in \M$: for every commutative square as the one below, with $e \in \E$ and $m \in \M$,
			there exists a unique $k: Y \to Z$ whitch fits in it.
			\[\xymatrix{
					X\ar[r]^{g} \ar[d]_e & Z \ar[d]^m \\ Y \ar@{.>}[ur]^k \ar[r]_f & V
			}\]
	\end{enumerate}
	A factorization system $(\E, \M)$ is \emph{proper} if every $e \in \E$ is epi and every $m \in \M$ is mono, while it is \emph{stable} is $\E$ is stable under pullbacks.
\end{definition}

\begin{remark}\label{rem:dual_fact_sys}
	It is immediate to notice that, given a factorization system $(\E, \M)$ on a category $\X$, then $(\M, \E)$ is a factorization system on $\X^{op}$, which is proper if $(\E, \M)$ is so, and it is stable if
	$\M$ is stable under pushouts in $\X$.
\end{remark}

\begin{remark}
	If $(\E, \M)$ is a factorization system for a category $\X$, then $\E$ is the class of all the arrows with the left lifting property with respect to every $m\in \M$.
	Indeed, if $f:X \to Y$ is an arrow with such a property with respect to every element of $\M$, it is possible to factorize it as $f = m \circ e$, with $m: M \to Y$ in $\M$ and 
	$e: X \to M$ in $\E$, obtaining the above situation by definition.
	\[\xymatrix{X \ar[r]^e\ar[d]_f & M \ar[d]^m \\ Y \ar[r]_{\id{Y}} \ar@{.>}[ur]^k & Y}\]
	From this, it is possible to prove that $k$ and $m$ are mutually inverses, hence $f = m\circ e$ belongs to $\E$ due to closure under isomorphisms.
\end{remark}

\begin{proposition}
	Let $(\E, \M)$ be a factorization system on $\X$. Then $\E$ contains every isomorphism, and it is closed under composition.
\end{proposition}

By \Cref{rem:dual_fact_sys}, we have also that $\M$ is closed under composition, and contains every isomorphism.

\begin{lemma}
	Let $(\E, \M)$ be a factorization system on $\X$. If $e:Y \to E$, $e': Y\to E'$ and $m: E \to X$, $m':E'\to X$ are arrows, respectively, in $\E$ and in $\M$ such that $e'\circ m' = e \circ m$,
	then there exists a unique isomorphism $\phi: E \to E'$ such that the diagram below commutes.
	\[\xymatrix{Y\ar[r]^{e'}\ar[d]_{e}&E'\ar[d]^{m'}\\E\ar[r]_{m}\ar@{.>}[ur]^{\phi}\ar[r]_{m}&X}\]
\end{lemma}

\begin{proof}
	Suppose the following situation.
	\[\xymatrix{X \ar[r]^e\ar[d]_{e'}&E\ar[d]^{m}&&X\ar[r]^{e'}\ar[d]_{e}&{E'}\ar[d]^{m'}\\{E'}\ar[r]_{m'}\ar@{.>}[ur]^{\phi}&Y&&E\ar[r]_{m}\ar@{.>}[ur]^{\psi}&Y}\]
	From the commutativity of the diagrams below, we obtain at once $\phi\circ\psi=\id{E'}$ and $\psi\circ\phi=\id{E}$.
	\[\xymatrix{
			X\ar[rr]^{e}\ar[dd]_{e}\ar[dr]_{e'}&&E\ar[r]^{m}\ar[dl]^{\psi}&Y\ar[dd]^{\id{Y}} &&& X\ar[rr]^{e'}\ar[dd]_{e'}\ar[dr]_{e}&&{E'}\ar[r]^{m}\ar[dl]^{\phi}&Y\ar[dd]^{\id{Y}}\\
							     & {E'}\ar[dl]_{\phi}\ar[drr]^{m'}& & &&&  & {E'}\ar[dl]_{\psi}\ar[drr]^{m'}& & \\
			E\ar[rrr]_{m} &&& Y &&& {E'} \ar[rrr]_{m'} &&& Y
	}\]
\end{proof}

The previous lemma allows us to deduce other useful properties of factorization systems.

\begin{proposition}
	Given a factorization system $(\E, \M)$ on a category $\X$, it holds that:
	\begin{enumerate}
		\item an arrow $f:X\to Y$ that factors as $m \circ e$ belongs to $\M$ if and only if $e$ is an isomorphism;
		\item $f\in \E$ and $f\in\M$ if and only if $f$ is an isomorphism;
		\item $\M$ is stable under pullbacks;
		\item if $(\E, \M)$ is proper, then $g\circ f \in \M$ implies $f \in \M$.
	\end{enumerate}
\end{proposition}

\todo{proof?}