### 2025-02-18 calco/Labeled_hyp_with_eq.tex

\subsection{Labeled Hypergraphs with Equivalences}

As we have done in \Cref{sssect:hyp_alg_sign}, we can define the category of hypergraphs with equivqlence labeled over an algebraic signature.

\begin{definition}
	Let $\Sigma = (O_\Sigma, \ari_\Sigma)$ be an algebraic signature, and let $\mathcal{G}^{\Sigma}$ the hypergraph associated to $\Sigma$.
	Then, the \emph{hypergraph with equivalence} associated to $\Sigma$ is $L(\mathcal{G}^\Sigma)$, and the category of hypergraphs with equivalence labeled over $\Sigma$ is the 
	slice category $\EqHyp_\Sigma := \EqHyp/L(\mathcal{G}^\Sigma)$.
\end{definition}

By \Cref{thm:slice_functors}, we can deduce the following.

\begin{proposition}
	$\EqHyp_\Sigma$ is $\reg (\EqHyp_\Sigma)$-adhesive.
\end{proposition}