### 2025-02-18 calco/eggs.tex

\section{EGGs}

\todo{introduction}

\begin{definition}
	Let $\mathcal{G} = (E, V, C, s, t, q)$ be a hypergraph with equivalence, and let $(S, \pi_1, \pi_2)$ be the kernel pair of $q^\star \circ s$.
	Then, $\mathcal{G}$ is an \emph{e-hypergraph} whenever $q^\star \circ t \circ \pi_1 = q^\star \circ t \circ \pi_2$.
	$\egg$ is the full subcategory of $\EqHyp$ where objects are e-hypergraphs, and $I: \egg \to \EqHyp$ is the inclusion functor.
\end{definition}

\begin{lemma}
	$\egg$ has all limits, and $I$ preserves them.
\end{lemma}

\beign{proof}

\end{proof}