### 2025-02-23 calco/Equivalence_term_graphs.tex
minima

\subsubsection{Term Graphs with Equivalences}

\begin{definition}
	Let $\Sigma$ be an algebraic signature.
	A labeled hypergraph with equivalence $l: \mathcal{G} \to L(\mathcal{G}^\Sigma)$ is a \emph{term graph with equivalence} if $t_{\mathcal{G}}$ is mono.
	We define category of term graphs with equivalence over $\Sigma$, denoted $\EqTG_{\Sigma}$, as the full subcategory of $\EqHyp_\Sigma$,
	and the corresponding inclusion functor $I_{\EqTG_\Sigma}$.
\end{definition}

\begin{proposition}
	If $l: \mathcal{G \to G}^\Sigma$ is a term graph, then $L(l)$ is a term graph with equivalence.
\end{proposition}