### 2025-02-23 calco/Labeled_hyp_with_eq.tex
ritocchi

\subsection{Labeled Hypergraphs with Equivalences}

As we have done in \Cref{sssect:hyp_alg_sign}, we can define the category of hypergraphs with equivalence labeled over an algebraic signature.

\begin{definition}
	Let $\Sigma = (O_\Sigma, \ari_\Sigma)$ be an algebraic signature and $\mathcal{G}^{\Sigma}$ the hypergraph associated to $\Sigma$.
	Then the \emph{hypergraph with equivalence} associated to $\Sigma$ is $L(\mathcal{G}^\Sigma)$ and the category of hypergraphs with 
	equivalence labeled over $\Sigma$ is the 
	slice category $\EqHyp_\Sigma = \EqHyp/L(\mathcal{G}^\Sigma)$.
\end{definition}

By \Cref{thm:slice_functors}, we can deduce the following.

\begin{proposition}
	$\EqHyp_\Sigma$ is $\reg (\EqHyp_\Sigma)$-adhesive.
\end{proposition}

We can lift the adjunction given by $T$ and $L$ to $EqHyp_\Sigma$ and $\hyp_\Sigma$.\todo{Non capisco se è uno svarione mio o ha senso}

By \Cref{cor:mono1}, we can deduce what follows.

\begin{proposition}\label{prop:monos_in_eqhyps}
	A morphism $h$ between two objects of $\EqHyps$ is mono if and only if $T(h)$ is mono.
\end{proposition}