### 2025-02-26 calco/Equivalence_term_graphs.tex
minima

\subsubsection{Term Graphs with Equivalences}

\begin{definition}
	Let $\Sigma$ be an algebraic signature.
	A labeled hypergraph with equivalence $l: \mathcal{G} \to R(\mathcal{G}^\Sigma)$ is a \emph{term graph with equivalence} if $t_{\mathcal{G}}$ is mono.
	We define category of term graphs with equivalence over $\Sigma$, denoted $\EqTG_{\Sigma}$, as the full subcategory of $\EqHyp_\Sigma$,
	and the corresponding inclusion functor $I_{\EqTG_\Sigma}$.
\end{definition}

\begin{proposition}
	If $l: \mathcal{G \to G}^\Sigma$ is a term graph, then $R(l)$ is a term graph with equivalence.
\end{proposition}

We have a functor $T_\EqTGs : \EqTGs \to \tg$, defined to be $T_\Sigma \circ I_\EqTGs$. Since both $T_\Sigma$ and $I_\EqTGs$ are faithful, so is $T_\EqTGs$.

If $h: \mathcal{H \to G}$ is a mono between $l': \mathcal{H} \to L(\mathcal{G}^\Sigma)$ and $l: \mathcal{G}\to L(\mathcal{G}^\Sigma)$, then, if $l\in\EqTGs$, $l'\in\EqTGs$ as well. This let us deduce what follows.

\begin{proposition}
	$\EqTGs$ has equalizers, and $I_{\EqTGs}$ creates them.
\end{proposition}