### 2025-05-05 calco/rebuttal/letter.tex
rebuttal

\documentclass[english,11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{cleveref}
\usepackage{fullpage}

\title{``EGGs are adhesive!''\\
	Answer to reviewers}
\author{R. Biondo, D. Castelnovo, F. Gadducci}

\begin{document}
	\maketitle

We wish to thank the reviewers for the time they have dedicated to our paper, and for the many useful suggestions. We corrected all typos indicated; below we explain how we have addressed the issues raised by the reviewers.

\section*{Reviewer 1}
\subsection*{Comments}
\begin{itemize}
\item  
\end{itemize}
\subsection*{Bibliography}
\begin{itemize}
	\item 
\end{itemize}

\subsection*{Sources of theorems and definitions}
\begin{itemize}
	\item 
\end{itemize}

\section*{Reviewer 2}

\end{document}